|

### 2019-09-30 analysis/paper/manuscript.Rmd
Address Keith's suggestions down to the Tagging section

Quantitative stock assessment plays a central role in modern fisheries management [@hilborn1992]. Over time, and as new methods are developed, there is an increasing amount of data available to inform stock assessments. These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series. Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2013]. For well-monitored stocks, the challenge has shifted from having sufficient data and information for providing sound advice on stock status, to presenting large quantities of data and output from increasingly complex statistical models in a meaningful way. Traditional formats (e.g. slides) for presenting information at stakeholder meetings are often insufficient to convey the richness of information available and the static and sequential nature of these formats can stifle meaningful discussions. A solution to this problem is the effective use of interactive visualization tools [@keena2016]. These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community. As a result, an increasing number of scientists across a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [e.g. @jones2016; @yeatman2018; @letcher2018]. This trend includes fisheries research and many of these tools, and the code used to produce them, are available online (see Appendix A for examples).

Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever-expanding array of data inputs and model outputs. First, we focus on data that are commonly presented in stock assessments and use interactive plots to simplify the detailed exploration of data from long-term monitoring programs program. Second, we focus on the modeling aspect of stock assessment and demonstrate how dynamic and interactive data visualizations can be used to explore, diagnose and communicate results from an integrated assessment model. This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling.

@@

<!-- should the concept of a dashboard be introduced here? -->

@@

Fisheries and Oceans Canada (DFO) has been conducting a multi-species stratified-random survey across the Newfoundland and Labrador shelf since the 1970s [@rideout2018]. Currently, these data are analyzed using a standard stratified analysis via a locally developed R package called RStrap [for details on methodology see @smith1981]. Both the inputs and outputs from the analysis are very large and, depending on the species, the time-series may include > 40 years and span the majority of the Newfoundland and Labrador shelf. In order for people to quickly and reliability explore these data without iteratively modifying R scripts, an app was built using a combination of shiny [@chang2018a], crosstalk [@cheng2016], flexdashboard [@allaire2017], plotly [@sievert2018], and R Markdown [@allaire2018] all within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio]. The app started as a way to visualize estimates of biomass and abundance trends of specific species using the flexdashboard package, however, shiny was latter applied as it allows stratified analyses to be run in the background and therefore allows the user to dynamically explore estimates from multiple species in the one session. Both crosstalk and plotly were incorporated to allow the user to interact with the data. The app contains four primary tabs ("Survey Indices", "Age & Length Distributions", "Recruitment", and "Help") containing stock level estimates of biomass (Figure \@ref(fig:biomass)) and abundance, mean numbers and weights per standardized survey tow, stock level biomass broken out by division, age and length frequencies (Figure \@ref(fig:freq)), and recruitment indices. The number of potential inputs to RStrap are rather large and most do not need to be modified. Some inputs however, do require modification. The input interface on the RStrap Explorer app is divided into two parts: basic and advanced inputs. Basic inputs are those like species, season, and survey year(s). Advanced inputs allow the user to select unconverted data (default) or converted data (which allows for a single complete time-series to be used as opposed to time-series broken by changes in survey gear) and select whether they would like analyses conducted by length and/or age, among others. This app was built for data exploration and allow the user to obtain quick and easy visualization of survey data, but is not meant to replace the through data checking and analysis required for formal stock assessment.

```{r biomass, echo = FALSE, fig.cap = '- Screen shot of RStrap Explorer tool, using flexdashboard, plotly, and shiny. On the left side of the figure, one can see the "Basic Inputs" sidebar, which is fixed across all pages and allows the user to specify the species, NAFO Divisions, season, and survey years of interest. The "Biomass Index" sub-tab is illustrated here. This sub-tab shows the unconverted time-series of biomass data with 95% confidence intervals (Engle and Campelen) for both cod (*Gadus morhua*; species code 438) and American Plaice (*Hippoglossoides platessoides*; species code 889). All trend lines can be toggled off using plotly’s dynamic user interface. The user also has the ability to hover over the points to compare estimates and export the figure as a png.'}
knitr::include_graphics("figures/RStrap_dashboard_biomass.png")
```

@@

In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive and self-contained "dashboard" was developed for the 2018 assessment of Northern cod [@dwyer2019]. The concept of using a dashboard was borrowed from the business community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators. We used R-based packages [@R] and the RStudio IDE [@Rstudio] to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard [@allaire2017] package to group interactive plotly-based [@sievert2018] visuals into a dynamic document. We also used the crosstalk [@cheng2016] package to link the data displayed across multiple plots. Via R Markdown [@allaire2018], the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data.

@@

- Marine Aggregates Application (https://openscience.cefas.co.uk/ma_tool/) from Cefas
- Benthic Non Native Species Tool (https://openscience.cefas.co.uk/invasive_species/) from Cefas
- Species Dashboard (https://shiny.marine.ie/speciesdash/) from The Marine Institute
- The Stock Book (https://shiny.marine.ie/stockbook/) from The Marine Institute
- Catchy Data (https://eu.oceana.org/en/catchy-data) from Oceana
- VISA tool (e.g. https://ices-tools-dev.github.io/VISA_tool/hke.27.3a46-8abd.html#ices_advice_2018) from ICES, which was developed to meet an "EU request on dissemination of ICES advice beyond pdf files"

>>>>

Quantitative stock assessment plays a central role in modern fisheries management [@hilborn1992]. Over time, and as new methods are developed, there is an increasing amount of data available to inform stock assessments. These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series. Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2013]. For well-monitored stocks, the challenge has shifted from having sufficient data and information for providing sound advice on stock status, to presenting large quantities of data and output from increasingly complex statistical models in a meaningful way. Traditional formats (e.g. slides) for presenting information at stakeholder meetings are often insufficient to convey the richness of information available and the static and sequential nature of these formats can stifle meaningful discussions. A solution to this problem is the effective use of interactive visualization tools [@keena2016]. These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community [@perkel2018]. As a result, an increasing number of scientists across a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [e.g. @jones2016; @yeatman2018; @letcher2018]. This trend includes fisheries research and many of these tools, and the code used to produce them, are available online (see Appendix A for examples).

Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever-expanding array of data inputs and model outputs. We especially highlight the use of interactive visuals inside dynamic documents called "dasbhoards". The concept of using a dashboard was borrowed from the business community where they are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators. We first demonstrate the application of these tools to data that are commonly presented in stock assessments to show how they can simplify the detailed exploration of data from long-term monitoring programs. Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive dasbhoards can be used to explore, diagnose and communicate results from an integrated assessment model. This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling. In each case, the dashboards were constructed within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio] using three or more of the packages listed in Table (Table \@ref(tab:pkg)).

```{r pkg, results = "asis", echo = FALSE}

## Package name and description
pkg <- c(shiny = 'Makes it incredibly easy to build interactive web applications with R. Automatic "reactive" binding between inputs and outputs and extensive prebuilt widgets make it possible to build beautiful, responsive, and powerful applications with minimal effort [@chang2018a]',
         shinydashboard = "Create dashboards with 'Shiny'. This package provides a theme on top of 'Shiny', making it easy to create attractive dashboards [@chang2018b].",
         rmarkdown = "Convert R Markdown documents into a variety of formats [@allaire2018].",
         flexdashboard = "Format for converting an R Markdown document to a grid oriented dashboard. The dashboard flexibly adapts the size of it's components to the containing web page [@allaire2017].",
         plotly = "Easily translate 'ggplot2' graphs to an interactive web-based version and/or create custom web-based visualizations directly from R [@sievert2018].",
         leaflet = "Create and customize interactive maps using the 'Leaflet' JavaScript library and the 'htmlwidgets' package. These maps can be used directly from the R console, from 'RStudio', in Shiny applications and R [@cheng2018].",
         crosstalk = 'Provides building blocks for allowing HTML widgets to communicate with each other, with Shiny or without (i.e. static .html files) [@cheng2016].')

tab <- data.frame(names(pkg), unname(pkg))
names(tab) <- NULL
# names(tab) <- c("**Package**", "**Description**")
tab[, 1] <- paste0("**`", tab[, 1], "`**")

knitr::kable(tab, caption = "Names and descriptions of the key packages used to build the interactive applications presented in this paper. Descriptions are directly from the `DESCRIPTION` file of each package.")

```

@@

@@

Fisheries and Oceans Canada (DFO) has been conducting a multi-species stratified-random survey across the Newfoundland and Labrador shelf since the 1970s [@rideout2018]. These data are analyzed using a standard stratified analysis via a locally developed R package called RStrap [for details on methodology see @smith1981]. Both the inputs and outputs from RStrap analyses are very large and, depending on the species, the time-series may include > 40 years and span the majority of the Newfoundland and Labrador shelf. In order for people to quickly and reliably explore these data without iteratively modifying R scripts one species at a time, an application (hereafter called RStrap Explorer) was built using a combination of shiny, crosstalk, flexdashboard, plotly and R markdown (Table \@ref(tab:pkg)). RStrap Explorer started as a way to visualize estimates of biomass and abundance trends of specific species by supplying RStrap output to a flexdashboard file designed to organize and visualize the results. The shiny package was latter applied as it allows RStrap to be run in the background and therefore allows the user to dynamically explore estimates from multiple species in the one session. Both crosstalk and plotly were incorporated to allow the user to interact with the data.

RStrap Explorer contains four primary tabs: 1) "Survey Indices" contains stock level estimates of biomass and abundance (Figure \@ref(fig:biomass)), 2) "Age & Length Distributions" contains length and age frequency plots (Figure \@ref(fig:freq)), 3) "Recruitment" displays recruitment indices, and 4) "Help" provides additional context to the survey data and analysis. The number of potential inputs to RStrap are rather large and most do not need to be modified. Some inputs however, do require modification. The input interface for RStrap Explorer is divided into two parts: basic and advanced inputs. Basic inputs are those like species, season, and survey year(s). Advanced inputs allow the user to define the length or age of recruitment, select whether they would like analyses conducted by length and/or age, and so on. This application was built for data exploration and allow the user to obtain quick and easy visualization of survey data, but is not meant to replace the through data checking and analysis required for formal stock assessment.

```{r biomass, echo = FALSE, fig.cap = '- Screen shot of the "Survey Indices" page of the RStrap Explorer application. The "Basic Inputs" sidebar is fixed across all pages and allows the user to specify the species, NAFO Divisions, season, and survey years of interest. The "Biomass Index" sub-tab is illustrated here. This sub-tab shows a time-series of biomass data with 95% confidence intervals, with a break in the lines when the survey gear changed from an Engle to Campelen trawl, for both cod (*Gadus morhua*; species code 438) and American Plaice (*Hippoglossoides platessoides*; species code 889). All trend lines can be toggled off using plotly’s dynamic user interface. The user also has the ability to hover over the points to compare estimates and export the figure as a png file.'}
knitr::include_graphics("figures/RStrap_dashboard_biomass.png")
```

@@

In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive and self-contained "dashboard" was developed for the 2018 assessment of Northern cod [@dwyer2019]. We used R-based packages [@R] and the RStudio IDE [@Rstudio] to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard [@allaire2017] package to group interactive plotly-based [@sievert2018] visuals into a dynamic document. We also used the crosstalk [@cheng2016] package to link the data displayed across multiple plots. Via R Markdown [@allaire2018], the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data.

@@

- Marine Aggregates Application (https://openscience.cefas.co.uk/ma_tool/) from Cefas
- Benthic Non Native Species Tool (https://openscience.cefas.co.uk/invasive_species/) from Cefas
- Species Dashboard (https://shiny.marine.ie/speciesdash/) from The Marine Institute
- The Stock Book (https://shiny.marine.ie/stockbook/) from The Marine Institute
- Data-Limited Fisheries Toolkit demo (https://www.datalimitedtoolkit.org/demo) from the Fisheries Center, University of British Columbia
- Exploring Groundfish in an Ecosystem Context (https://incorporatingecosystemapproach.shinyapps.io/indiapp/) from Fisheries and Oceans Canada, Maritimes Region
- Catchy Data (https://eu.oceana.org/en/catchy-data) from Oceana
- VISA tool (e.g. https://ices-tools-dev.github.io/VISA_tool/hke.27.3a46-8abd.html#ices_advice_2018) from ICES, which was developed to meet an "EU request on dissemination of ICES advice beyond pdf files"
